Elliptic operator stability and small cells on finite volume grids

### 2023-05-19 example/EBCM/hoeb_petsc/_doc/paper.tex
good progress on document

  Numerical stability for elliptic operators is well understood.  To
  be numerically stable, a numerical operator for an elliptic equation
  cannot have non-zero positive real eigenvalues.  We explore a family
  of high order finite volume elliptic operators. We attempt to
  define for this family of algorithms the boundaries of stability
  within its parameter space.

@@

Given a sufficiently smooth function $\psi$, we can approximate $\psi$
in the neighborhood of $\xbar$ using a Taylor expansion to order $P_T$:
\begin{equation}
\psi(\xbold)  =  \sum\limits_{p < P_T} C^p (\xbar -\xbar)^p
\label{eqn::taylor}
\end{equation}
where $C^p$ this appropriate Taylor coefficient.  In three dimensions,
\begin{equation}
  C^p =\frac{1}{p!}
      \frac{\partial^{p_0}}{\partial x_0}
      \frac{\partial^{p_1}}{\partial x_1}
      \frac{\partial^{p_2}}{\partial x_2}  (\psi).
\end{equation}  
In finite volume  methods, we define grid data to be
averages over volumes.   So if we discretize the smooth function
$\psi$, the average over the volume $\ibold$ is given by
\begin{equation*}
 V_\ibold <\psi>_\ibold = \int\limits_{V_\ibold} \psi(\xbold) dV
\end{equation*}
If we insert the Taylor expansion \ref{eqn::taylor}, we get a discrete
approximation to the smooth function $\psi$ that is accurate to order $P_T$.
\begin{equation*}
  <\psi>_\ibold = \frac{1}{V_\ibold} \sum\limits_{p < P_T} C^p m^p
  \label{eqn::vol}.
\end{equation*}
This forms a local polynomial expansion of $\psi$ around the volume
$\ibold$ expressed in terms moments ($m$).
Modern, higher order embedded boundary
methods use this description directly to generate stencil coefficients
\cite{Overton2022a, Devendran2017, Schwartz2015, Katz2023}.

@@

Devendran, et al. \cite{Devendran2017} present an algorithm that
solves Poisson's equation to fourth-order accuracy using an EB grid.
Generalizing somewhat  the framework they present, one can define an
elliptic operator of any accuracy.

At each volume $V$, we define the neighborhood $\neigh_v$
volumes within a radius $R_s$ number of cells from the volume.
Fix $\xbar$ to be the distance from the center of the target cell.
Let's say there are
$N_v$ volumes in the neighborhood and $F_D$ domain faces and $F_C$ cut
cell faces in the domain.  Following (\cite{Devendran2017}), we create
an equation for each volume and boundary condition within the
neighborhood to form an  overdetermined system of equation for the
Taylor coefficients in $V$.    Each volume gets an equation of the
form of equation \ref{eqn::vol}.    Faces at the boundary domain with
Neumann boundary conditions get an equation of the form 
by We For each volume, we have The equations for the volumes and the
boundary conditions form an overdetermined  system.

>>>>

Devendran, et al. \cite{Devendran2017} present an algorithm that
solves Poisson's equation to fourth-order accuracy using an EB grid.
Generalizing somewhat  the framework they present, we can define a
family of elliptic operators.
Here we investigate how the accuracy and stability of this family
of operators varies within its parameter space.

@@

@@

Devendran, et al. \cite{Devendran2017} present an algorithm that
solves Poisson's equation to fourth-order accuracy using embedded
boundaries.  Generalizing somewhat  the framework they present, we
present a family of elliptic operators.

We are solving Poisson's equation for field $\psi$, given charge $\rho$:
\begin{equation}
  \nabla \cdot (\nabla \psi) = \rho
  \label{eqn::poisson}
\end{equation}
If we integrate equation \ref{eqn::poisson} and apply the divergence theorem 
over control volumes we get
\begin{equation}
\int_{V_\ibold} \nabla \cdot (\nabla \phi ) dV = \int_{\partial
  V_\ibold} \nabla \phi\cdot \nhat  dA = \int_{V_\ibold} \rho dV
\end{equation}
Without approximation, we can say that the surface integral can be
computed by a sum over the faces that comprise the surface of the
volume.
\begin{equation}
\sum_{f: f \in \partial V_\ibold} \nabla \phi \cdot \nhat dA = \int_{V_\ibold} \rho dV
\end{equation}
Using equations \ref{eqn::faceavg} and \ref{eqn::volavg}, this becomes 
\begin{equation}
\sum_{f: f \in \partial V_\ibold} \avg{\nabla \phi \cdot \nhat}_f A_f
= \avg{\rho}_\ibold V_\ibold
\label{eqn::discretePoisson}
\end{equation}
A finite volume operator is defined by the algorithm to
compute the integral of  fluxes at each face.   We have three types of
face, open faces (the faces between two volumes that are aligned with
coordinate directions), domain  faces, and cut faces.    For now, let's say we have
homogeneous Neumann boundary conditions at domain and cut faces $\nabla \psi
\cdot \nhat  = 0 |\partial \Omega$.   In this case, to describe the
algorithm, we need only describe how to compute $<\nabla \psi \cdot
\nhat>$ on open faces.

\section{Open face flux}
\section{Dirichlet fluxes at boundaries}
We can approximate $\psi$
in the neighborhood of $\xbar$ using a Taylor expansion to order $P_T$:
\begin{equation}
\psi(\xbold)  =  \sum\limits_{p < P_T} C^p (\xbar -\xbar)^p
\label{eqn::taylor}
\end{equation}
where $C^p$ this appropriate Taylor coefficient.  In three dimensions,
\begin{equation}
  C^p =\frac{1}{p!}
      \frac{\partial^{p_0}}{\partial x_0}
      \frac{\partial^{p_1}}{\partial x_1}
      \frac{\partial^{p_2}}{\partial x_2}  (\psi).
\end{equation}
If we know the local Taylor coeffiencts,
the gradient of the field  in direction $d$ is given by
\begin{equation}
  \frac{\partial \psi}{\partial x^d} \psi =
  \sum\limits_{p < P_T} p^d C^{p-e^d} (\xbar -\xbar)^{p-e^d}
\end{equation}

In finite volume  methods, we define grid data to be
averages over volumes and fluxes between volumes are averages over the
faces between volumes.   
For the field $\psi$, the average over the volume $\ibold$ is given by
\begin{equation*}
 V_\ibold <\psi>_\ibold = \int\limits_{V_\ibold} \psi(\xbold) dV
\end{equation*}
If we insert the Taylor expansion \ref{eqn::taylor}, we get a discrete
approximation to the smooth function $\psi$ that is accurate to order $P_T$.
\begin{equation*}
  <\psi>_\ibold = \frac{1}{V_\ibold} \sum\limits_{p < P_T} C^p m^p
  \label{eqn::vol}.
\end{equation*}
Similarly, integrating the flux over a d-directional face $\fbold$:
\begin{equation*}
  <\nabla \psi_d>_\ibold = \frac{1}{A_f\bold}
  \sum\limits_{p < P_T} p^d C^{p-e^d} m^{p-e^d}
  \label{eqn::vol}.
\end{equation*}
To solve for the Taylor coefficients, we use the neighboring volumes's
values  to form a system of equations (equation \eqn::vol
for each volume $\ibold \in N_f$, the neighborhood of $f$).  Boundary
condition equations are also added to the system.   We force there
to be enough neighbors so that this system is overdetermined.

\section{}
